Parallelization of Fast $\ell_1$-Minimization for Face Recognition

### 2011-05-28 main.tex
Finished first pass through Allen's notes. Still have critical TODO items

First, in order to test our theory that the GPU performs best solving many
$\ell_1$ problems-per concurrently, we benchmark three implementations of the
alignment $\ell_1$-min with $A$ of size $5120 \times 32$, and a fixed
$50$ inner loop iterations for each of $50$ outer loop iterations. The runtime
on a GTX480 GPU is averaged over a large number of trials, which are run
sequentially or concurrently depending on the implementation.
The results are shown in Table \ref{tbl:ubench}.
Our proposed parallelization of the $\ell_1$-min used in the alignment
stage is {\em eight times} faster than an implementation solving a single
problem at a time using the stock BLAS libraries.
%The results are shown below:
\begin{table}[h!]
\caption{Benchmark of implementations of alignment $\ell_1$-min.}
\begin{tabular}{|l|c|}
\hline
Sequential solver using CUBLAS & 302\,ms \\
\hline
One problem per SMP using CUDA streams & 70\,ms  \\
\hline
One problem per SMP using single kernel & {\bf 36\,ms} \\
\hline
\end{tabular}
\label{tbl:ubench}
\end{table}
%\footnote{The streams implementation is limited significantly by a 
%cap on the number of concurrent streams in the current version of CUDA}

For experiments on face data we use subsets of the CMU Multi-PIE Face Database.
For gallery images we use frontal images from session 1, which contains 20
images per user taken under different illuminations. For test images we use
images from session 2.  The training images are prepared as follows:
The iterative alignment stage seeks a similarity
transformation between the coordinate frame of the "camera" resolution test
image and the coordinate frame of the "canonical" resolution images which the
images are compared.  We use a $64 \times 64$ pixel window in the canonical
frame.  The training images are aligned by manually clicking the two outer eye
corners in the camera frame, and then warping the images via a similarity
transformation such that the eye corners lie at a pair of canonical eye
locations in the canonical frame.  The canonical eye corners are chosen such
that for well aligned images, the canonical window contains exclusively pixels
from the subject's face.  We assume the face detector is accurate, so for the test
images we use clicked eye corners for test image alignment initialization.\footnote{Previous
publications have shown that iterative alignment improves recognition accuracy substantially
even with clicked eye corners}

In Figure \ref{fig:alignment_stage_runtime} we report the time it takes fort
our proposed CPU and GPU algorithms to align a single test image against each
of the users in the gallery.  We vary the size of the gallery to show how
implementation scales in the number of gallery users.
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/alignment_runtime_graph}
\caption{Alignment stage runtime vs. size of training database.}
\label{fig:alignment_stage_runtime}
\end{figure}
Similar to the results in Figure \ref{fig:random_data},
there is a crossover point above which the GPU out-performs the CPU.  Since
the CPU algorithm solves problems sequentially, there is a linear trend.  
The tiers seen in the GPU profile result from the GPU hardware scheduling alignment
problems in concurrent batches, but the overall trend is linear as expected.

@@

Finally, we perform an experiment verifying the recognition accuracy of the
overall pipeline.  Since would expect higher recognition rates but longer
execution time for higher resolution images, we compare the recognition rate of
the overall pipeline with the face recognition stage run at 5 different
canonical image resolutions: $32\times32, 48 \times 48, 64 \times 64, 96 \times
96, \textrm{and} 128 \times 128$.  As motivated by Figure
\ref{fig:user_alignment_rank_plot}, we fix the number of gallery users to 20,
giving approximately 95\% probability to have the true user in the training
set.  
%The inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.  
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/accuracyVsResolution}
\caption{Recognition rate of the full pipeline vs. face window resolution.}
\label{fig:accuracy_vs_resolution}
\end{figure}
As can be seen in Figure \ref{fig:accuracy_vs_resolution}, at the optimal
resolution, which happens to match the alignment stage resolution, this bound
is very nearly achieved by the GPU implementation.  For significantly lower
resolutions, the accuracy drops off significantly.  Of the two main stages of
the face recognition pipeline, the iterative alignment stage is by far the most
expensive; the recognition stage requires less than one second.

>>>>

First, in order to test our theory that the GPU performs best solving many
$\ell_1$ problems-per concurrently, we benchmark three implementations of the
alignment $\ell_1$-min solver with $A$ of size $5120 \times 32$, and a fixed
$50$ inner loop iterations for each of $50$ outer loop iterations. The runtime
on a GTX480 GPU is averaged over a large number of trials, which are run
sequentially or concurrently depending on the implementation.
The results are shown in Table \ref{tbl:ubench}.
Our proposed parallelization of the $\ell_1$-min used in the alignment
stage is {\em eight times} faster than an implementation solving a single
problem at a time using the stock BLAS libraries.
\footnote{Note that the streams implementation is limited significantly by a 
cap on the number of concurrent streams in the current version of CUDA}
%The results are shown below:
\begin{table}[h!]
\caption{Benchmark of implementations of alignment $\ell_1$-min.}
\begin{tabular}{|l|c|}
\hline
Sequential solver using CUBLAS & 302\,ms \\
\hline
One problem per SMP using CUDA streams & 70\,ms  \\
\hline
Four problems per SMP using single kernel & {\bf 36\,ms} \\
\hline
\end{tabular}
\label{tbl:ubench}
\end{table}

Using an implementation motivated by the previous result, we now benchmark our
iterative alignment implementation on real face data.  For experiments on face
data we use subsets of the CMU Multi-PIE Face Database.  For gallery images we
use frontal images from session 1, which contains 20 images per user taken
under different illuminations. For test images we use images from session 2.
The training images are prepared as follows: The iterative alignment stage
seeks a similarity transformation between the coordinate frame of the "camera"
resolution test image and the coordinate frame of the "canonical" resolution
images which the images are compared.  We use a $64 \times 64$ pixel window in
the canonical frame.  The training images are aligned by manually clicking the
two outer eye corners in the camera frame, and then warping the images via a
similarity transformation such that the eye corners lie at a pair of canonical
eye locations in the canonical frame.  The canonical eye corners are chosen
such that for well aligned images, the canonical window contains exclusively
pixels from the subject's face.  We assume the face detector is accurate, so
for the test images we use clicked eye corners for test image alignment
initialization.\footnote{Previous publications have shown that iterative
alignment improves recognition accuracy substantially even with clicked eye
corners}

Figure \ref{fig:alignment_stage_runtime} shows the average run time of the CPU and GPU implementations to align
a query image against all the subject classes separately. We vary the number of
subject classes to show how the algorithms scale. 
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/alignment_runtime_graph}
\caption{Alignment stage runtime vs. size of training database.}
\label{fig:alignment_stage_runtime}
\end{figure}
Similar to the results in Figure \ref{fig:random_data},
there is a crossover point above which the GPU out-performs the CPU.  Since
the CPU algorithm solves problems sequentially, there is a linear trend.  
The tiers seen in the GPU profile result from the GPU hardware scheduling alignment
problems in concurrent batches, but the overall trend is linear as expected.

@@

Finally, we perform an experiment verifying the recognition accuracy of the
overall pipeline.  Since one would expect higher recognition rates but longer
execution time for higher resolution images, we compare the recognition rate of
the overall pipeline with the face recognition stage run at 5 different
canonical image resolutions: $32\times32, 48 \times 48, 64 \times 64, 96 \times
96, \textrm{and} 128 \times 128$.  As motivated by Figure
\ref{fig:user_alignment_rank_plot}, we fix the number of gallery users to 20,
giving approximately 95\% probability to have the true user in the training
set.  
%The inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.  
\begin{figure}
\centering
\includegraphics[width=3.4in]{figures/accuracyVsResolution}
\caption{Recognition rate of the full pipeline vs. face window resolution.}
\label{fig:accuracy_vs_resolution}
\end{figure}
As can be seen in Figure \ref{fig:accuracy_vs_resolution}, at the optimal
resolution, which happens to match the alignment stage resolution, this bound
is very nearly achieved by the GPU implementation.  For significantly lower
resolutions, the accuracy drops off significantly.  Of the two main stages of
the face recognition pipeline, the iterative alignment stage is by far the most
expensive; the recognition stage requires less than one second.
